## Loading and preprocessing the data
**Downloading and unzipping the dataset**
```{r}
library(dplyr)
library(magrittr)
url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url, destfile = "activity.zip", mode = "wb")
unzip("activity.zip")
```

**Reading the data from the dataset**
```{r}
activity = read.csv("activity.csv", header=T, colClasses = c("integer", "character", "integer") ,sep = ",")
```

**Converting character format to date format**
```{r}
activity$date <- as.Date(activity$date)
```

==================================================================================================
## What is mean total number of steps taken per day?
1. Get the total steps per day
```{r}
totalsteps <- aggregate(steps ~ date, activity, FUN = sum, na.rm = TRUE)
colnames(totalsteps) <- c("Days","Steps")
```

2. Create Histogram
```{r}
hist(totalsteps$Steps, main = paste("Total Steps Each Day"), col="red", xlab="Number of Steps")
```

3. Calculate mean and median
```{r}
mean_totalsteps <- mean(totalsteps$Steps)
median_totalsteps <- median(totalsteps$Steps)
```

The mean and median can be found in *mean_totalsteps* and *median_totalsteps* respectively.

=============================================
## What is the average daily activity pattern?
Step 1: Calculate the average steps for each interval for all days
```{r}
average_steps <- aggregate(steps ~ interval, activity, FUN = mean, na.rm = TRUE)
colnames(average_steps) <- c("Interval", "Steps")
```

Step 2: Plot the Average Number teps per day by Interval.
```{r}
plot(average_steps$Interval, average_steps$Steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
```

Step 3: Find interval with most average steps.
```{r}
max_numb_steps_interval <- average_steps[which.max(average_steps$Steps),1]
```

The average_steps contains the average steps for each interval for all days.

======================================
## Imputing missing values
A. Calculate the total number of missing values.
```{r}
sum(is.na(activity$steps))
```

B. Before creating a new dataset with missing data filled in, the missing values have to be imputed using mean for each day.
```{r}
activity2 <- subset(activity, !is.na(activity$steps))
affected <- is.na(activity$steps)
tAvg <- tapply(activity2$steps, activity2$interval, mean, na.rm = TRUE, simplify = T)
activity$steps[affected] <- tAvg[as.character(activity$interval[affected])]
```

C. Create a histogram of total  number of steps taken each day. Compare imputed histogram and NA histogram.
```{r}
totalsteps2 <- aggregate(steps ~ date, activity, FUN = sum)
par(mar = c(3,4,4,1)) 
hist(totalsteps2$steps, breaks = 10, col = "blue", xlab = "Daily Steps", main = "Total Steps Each Day")
hist(totalsteps$Steps, breaks = 10, col = "red", xlab = "Daily Steps", main = "Total Steps Each Day", add=T) 
legend("topright", c("Imputed", "NA"), fill = c("blue","red"), cex = 0.55)
```

D. Calculate the mean and median of the total number of steps taken per day.
```{r}
mean(totalsteps$Steps)
mean(totalsteps2$steps)

median(totalsteps$Steps)
median(totalsteps2$steps)
```

Based on the final findings, the means remain unchanged, however the median are different.

=====================================================
## Are there differences in activity patterns between weekdays and weekends?
(1) Create new factor variable in dataset with 2 levels - "weekday" and "weekend" to indicating if is a weekend or not.
```{r}
library(lubridate)
is_weekday <-function(date){
        if(wday(date)%in%c(1,7)) result<-"weekend"
        else
                result<-"weekday"
        result
}

activity_without_NAs <- mutate(activity,date=ymd(date)) %>% mutate(day=sapply(date,is_weekday))

table(activity_without_NAs$day)
```

(2) Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
library(ggplot2)
pattern <- activity_without_NAs %>% mutate(day=factor(day,levels=c("weekend","weekday")),steps_no_NAs=as.numeric(steps)) %>% group_by(interval,day) %>% summarise(average=mean(steps))
qplot(interval,average,data=pattern,geom="line",facets=day~.)
```

Based on the finding, on weekday basis, the test suject is more active at the start of the day, however, the activeness is higher throughout the weekends. 

